---
title: "cd-indicator-creation"
author: "Kelley Bemis"
date: "9/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(odbc)
library(DBI)
library(keyring)
library(dbplyr)

munis <- read_csv("https://github.com/Cook-County-Department-of-Public-Health/ccdph-data-sets/blob/main/2020/decennial-2020-total-muni.csv?raw=TRUE") %>%
  filter(district != "OOJ" & exclude_from_analysis == F) %>%
  mutate(geoid = paste0("17", census_place_code))

#function to censor data
suppress_count <- function(number_var) { number_var = ifelse(number_var > 0 & number_var < 5, NA, number_var) }

#function to calculate rates
rate <- function(num, pop, pop_denom, decimals = 1) { round(num/pop*pop_denom, decimals) }

```

## Package CD Indicators for SCC Health Atlas

Collect raw data from PH server

```{r collect-data}

#connect to server
cd_con <- dbConnect(odbc::odbc(), Driver = "SQL Server", Server = key_get("ph-server"), Database = "cd")

#pull CT and GC for year(s) of interest
sti_raw <- tbl(cd_con, "sti") %>%
  filter(year > 2016 & dx_code < 700) %>%
  collect()

##pull syphilis for year(s) of interest
syph_raw <-  tbl(cd_con, "sti") %>%
  filter(year > 2017 & dx_code %in% c(710,720)) %>%
  collect()

#disconnect from server
dbDisconnect(cd_con)

```

Create rates by muni, year, and disease

Note: Syphilis counts too suppressed for 3 year rates, will re-assess at 5 years

```{r create-rates}

#calculate annual STI rates
sti_rates <- sti_raw %>%
  mutate(topic = ifelse(dx_code < 300, "JG8FXNR", "JG8XQNO")) %>%
  group_by(year, clean_city, topic) %>%
  tally() %>%
  ungroup() %>%
  right_join(select(munis, municipality, pop_in_cook, geoid), by = c("clean_city" = "municipality")) %>%
  complete(clean_city, year, topic, fill = list(n = 0)) %>%
  group_by(clean_city) %>%
  fill(pop_in_cook, geoid) %>%
  ungroup() %>%
  mutate(n = suppress_count(n),
         rate = rate(n, pop_in_cook, 100000))

#calculate 3-year rates for syphilis (smaller numbers)
# syphilis_rates <- syph_raw %>%
#   mutate(period = "2018-2020") %>% #adjust cuts in the future when more data acquired
#   mutate(topic = "JG8ITHO") %>%
#   group_by(topic, period, clean_city) %>%
#   tally() %>%
#   ungroup() %>%
#   right_join(select(munis, municipality, pop_in_cook, geoid), by = c("clean_city" = "municipality")) %>%
#   complete(clean_city, period, topic, fill = list(n = 0)) %>%
#   group_by(clean_city) %>%
#   fill(pop_in_cook, geoid, .direction = "downup") %>%
#   ungroup() %>%
#   drop_na(period,topic) %>%
#   mutate(n = suppress_count(n),
#          rate = rate(n, pop_in_cook, 100000))
  

```

Format for Metopio upload

#Note: "NA"s must be removed manually for exported CSVs below before uploading to Metopio platform

```{r format-upload}

metopio_sti <- sti_rates %>%
  select(geoid, topic, year, rate) %>%
  pivot_wider(names_from = c(topic, year), values_from = rate, names_sep = "_") %>%
  drop_na(geoid) %>%
  mutate(layer = "place") %>%
  unnest(cols = contains("_"))

write_csv(metopio_sti, "cd/sti_2017-2020.csv")

# metopio_syph <- syphilis_rates %>%
#   select(geoid, topic, period, rate) %>%
#   pivot_wider(names_from = c(topic, period), values_from = rate, names_sep = "_") %>%
#   drop_na(geoid) %>%
#   mutate(layer = "place") %>%
#   unnest(cols = contains("_"))


#Note: remove exported "NA"s manually in CSV before uploading

```


## Package COVID Indicators for SCC Health Atlas

Note: Code below is temporary while datasets stored locally. Incorporate into above chunks when data moved to PH server

```{r covid}

covid_rates <- readRDS(paste0(key_get("local_shiny_path"), "cases.rds")) %>%
  mutate(year = lubridate::year(combinedReportDate)) %>%
  filter(year < lubridate::year(Sys.Date())) %>%
  group_by(year, clean_city) %>%
  tally() %>%
  ungroup() %>%
  right_join(select(munis, municipality, pop_in_cook, geoid), by = c("clean_city" = "municipality")) %>%
  complete(clean_city, year, fill = list(n = 0)) %>%
  group_by(clean_city) %>%
  fill(pop_in_cook, geoid) %>%
  ungroup() %>%
  mutate(n = suppress_count(n),
         rate = rate(n, pop_in_cook, 100000),
         topic = "JG8YTPN")

metopio_covid <- covid_rates %>%
  select(geoid, topic, year, rate) %>%
  pivot_wider(names_from = c(topic, year), values_from = rate, names_sep = "_") %>%
  drop_na(geoid) %>%
  mutate(layer = "place") %>%
  unnest(cols = contains("_"))

write_csv(metopio_covid, "cd/covid_2020-2021.csv")

```

